%!TEX root = Report.tex

\subsection{Related Work}

\noindent \textbf{TERMES: An Autonomous Robotic System for Three-Dimensional Collective Construction} \par

Paper by Kirstin Petersen, Radhika Nagpal and Justin Werfel \cite{termes1}. This paper presents a hardware system and high-level control scheme for autonomous construction of 3D structures. The hardware system consist of mobile autonomous robots and a collection of passive marked blocks. The general idea is to use swarm robotics to enable efficient construction even in extraterratial or disaster areas. The behavior of the robots is much like ants and termites in thaat sensing is entirely onbaord and each robot acts indenpendently of each other. Our contribution to the field can help enhance this project in the long run through analysis of the collaboration patterns of natures own swarms: ants and termites. By analysing how termites build their large hives one could potentially find a great way for swarm robots to collaborate when building structures. \\

\noindent \textbf{Distributed Multi-Robot Algorithms for the TERMES 3D Collective Construction System } \par

Paper by Kirstin Petersen, Radhika Nagpal and Justin Werfel \cite{termes2}. This paper present the algorithmic part of the TERMES 3D collective construction system described in the paper above \cite{termes1}. It describes algorithms for autonomous robot to build 3D structures both alone and in collaboration with more identical robots. Our contribution can potentially help enhance these algorithms. The movement of ants and termites could provide the key to better collaboration between the robots. Especially conflict resolution when two ants or termites need to occupy the same space on a path could be interesting to transfer into the robots' behavior.
